---
title: "p8105_hw2_xw2757"
output: github_document
---
```{r}
library(tidyverse)
library(readxl)
```
##Problem 1
**Loading and cleaning data**
```{r}
mtw_df <- 
  read_excel("TW_data.xlsx", sheet = "Mr. Trash Wheel", skip = 1, ) %>% 
  janitor::clean_names() %>% 
  filter(dumpster != "" , dumpster != "Grand Total") %>% 
  select(dumpster:homes_powered) %>% 
  mutate(month = factor(month), year = factor(year), dumpster = as.numeric(dumpster), sports_balls = round(sports_balls, digit = 0))
```



```{r}
p2019_df <- 
  read_excel("TW_data.xlsx", sheet = "2019 Precipitation", skip = 1, ) %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  mutate(year = "2019") %>% 
  rename(prcp = total)

p2018_df <- 
  read_excel("TW_data.xlsx", sheet = "2018 Precipitation", skip = 1, ) %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  mutate(year = "2018") %>% 
  rename(prcp = total)

prcp_df <-  
  bind_rows(p2018_df, p2019_df)

prcp_df %>% 
  mutate(month = month.name[month])
```

**Describing data**
In data set "Mr. Trash Wheel", the variables descirbed are as followed: `r names(mtw_df)`. There is a total number of `r nrow(mtw_df)` from the following years: `r levels(pull(mtw_df, year))`. Of all the trash categories, total number of sports balls picked up is `r sum(pull(mtw_df, sports_balls))`, the mean number is `r mean(pull(mtw_df, sports_balls))`, and the median is `r median(pull(mtw_df, sports_balls))`.


In precipitation data set, there are a total of `r nrow(prcp_df)` observations. The mean precipitation in 2018 is `r mean(pull(p2018_df, prcp))`, and the median is `r median(pull(p2018_df, prcp))`. The mean precipitation in 2019 is `r mean(pull(p2019_df, prcp))`, and the median is `r median(pull(p2019_df, prcp))`. Total precipitation in 2018 is `r sum(pull(p2018_df, prcp))`.

##Problem 2

**Importing and merging data**
```{r}
pols_month <- 
  read_csv(file = "fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, c("year", "month", "day"), sep = "-" , convert = TRUE) %>% 
  mutate(month = month.abb[month], 
         president = ifelse(prez_dem == 1, "dem", "gop")) %>% 
  select(-day, -prez_dem, -prez_gop)
         
snp <- 
  read_csv(file = "fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, c("month", "day", "year"), sep = "/") %>% 
  mutate(year = ifelse(year <= "15" ,  paste0("20", year), paste0("19", year)), month = as.numeric(month)) %>% 
  arrange(year, month) %>% 
  mutate(month = month.abb[month], year = as.numeric(year)) %>% 
  select(-day)
  

unemply <- 
  read_csv(file = "fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec, names_to = "month", values_to = "unemployment")
  
snp_pols <- 
  left_join(pols_month, snp, by = c("year" = "year" , "month" = "month"))

full_df <- 
  left_join(snp_pols, unemply, by = c("year" = "Year" , "month" = "month"))
```

The data set "full_df" describes informations compiled from data sets "snp", "unemply", and "pols_month". There are a total of `r nrow(full_df)` observations. "snp" contains information on S&P close number. "unemply" contains employment rate. "pols_month" contains information on the dem and gop number of representatives. The final data set contains data from year `r min(pull(full_df, year))` to `r max(pull(full_df, year))`. 


##Probelm 3
**Import data**
```{r}
baby_names <- 
  read_csv(file = "Popular_Baby_Names.csv", col_type = "fffcii") %>% 
  janitor::clean_names() %>% 
   mutate(ethnicity = replace(ethnicity, ethnicity == "ASIAN AND PACIFIC ISLANDER", "ASIAN AND PACI"),
         ethnicity = replace(ethnicity, ethnicity == "BLACK NON HISPANIC", "BLACK NON HISP"),
         ethnicity = replace(ethnicity, ethnicity == "WHITE NON HISPANIC", "WHITE NON HISP"),
         childs_first_name = str_to_lower(childs_first_name)) %>% 
  distinct(childs_first_name, year_of_birth,ethnicity, .keep_all = TRUE)

```
**Olivia Ranking**

```{r}
olivia_rank <- 
  filter(baby_names, childs_first_name == "olivia") %>%
  select(year_of_birth, gender, ethnicity, rank ) %>% 
  pivot_wider(names_from = "year_of_birth", values_from = "rank")

olivia_rank
```

**Male name over time**
```{r}
male_rank <- 
  filter(baby_names, gender == "MALE") %>% 
  select(year_of_birth, ethnicity, childs_first_name, rank) %>% 
  pivot_wider(names_from = "year_of_birth", values_from = "rank")
  
male_rank
```

**scatter plot**

```{r}

wtnonhisp_2016 <- 
  filter(baby_names, ethnicity == "WHITE NON HISP", gender == "MALE", year_of_birth == "2016") %>% 
  select(year_of_birth, childs_first_name, count, rank )

wtnonhisp_2016_plot <- 
wtnonhisp_2016 %>% 
  ggplot(aes(x = rank, y = count)) +
  geom_point(alpha = .5) + 
  labs(
    title = "Names of White Non Hispanic Male Born in 2016",
    x = "Rank in popularity of that name",
    y = "Number of children with a name"
  )

wtnonhisp_2016_plot
```

